Tabular Constraint Learning

### 2016-04-28 Paper/ECAI/tabular.tex
Merge

We consider various sets of cells in tables:
\begin{itemize}
  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
  \item
  A \textbf{vector} is either a column or a row that is type-consistent.
  If a vector is a row (column), we say that it has the \textit{row} (\textit{column}) orientation.
  \item
  A \textbf{group} is a subrange of vectors with the same orientation in a table.
  We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
  The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
  \item
 A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$. %It defines what it means for a constraint to be well-formed
%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
\end{itemize}
%For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}

@@

\paragraph{Example}
Several example constraints are illustrated in Table 1.  For example, the \template \textit{rank} has
\begin{itemize}
\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are the arguments;
\item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
of $v_x$ and $v_r$ be identical, i.e., $length(v_x) = length(v_r)$;
\item \CFunction: \textit{$v_r$ = RANK($v_x$)} whenever each value of $v_r$ is the rank of the corresponding value in $v_x$.
% is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
\end{itemize}

When looking into constraints like \textit{$v_r$ = RANK($v_x$)}, it is important to see the analogy with both first order logic and with constraint satisfaction problems.
From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $v_r$ and $v_x$ are the terms, which can be either variables or values.
This also holds in our setting: when we write
      $\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}$,
we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $[1,2,3,4]$. {\bf CHECK}
This also means that we can talk about constraints being satisfied. Basically, a constraint $c(t_1, ..., t_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
if the constraint holds on the values that the $t_i$ take in the set of tables ${\cal T}$. Figure 3 provides a list of constraints that are satisfied by the tables
in Figure 1.

@@

@@

Tables are detected by detecting rectangular collections of data and attempting to detect and separate headers from the data.
Especially this last step is a complex task on its own  \cite{header} and a sophisticated implementation goes beyond the scope of this paper.
On well formated data, however, the user will not have to specify tables manually.

Our algorithm also attempts to detect groups within tables by partitioning the data into consecutive, type consistent groups, essentially separating textual and numerical data.
We support some extensions such as currencies and percentual data and our approach allows us to omit manual specifications of groups.

@@

A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
We consider $c_1$ to be more specific than $c_2$ or $c_2$ to be more general than $c_1$.
A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.

@@

The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
\\\\
Most parts of the algorithm could be easily parallelized.
However, \dependencies would impose a partial order in which constraint templates would have to be treated.

@@

Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that those groups or some of their subgroups satisfy the \CSignature.

@@

\subsection{Case study}
Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
Figure~\ref{fig:sol_example} shows the constraints that we expect to find.

@@

\subsubsection*{Q2. How many redundant constraints are discovered?}
Our primary focus was accuracy, therefore we sometimes traded off less redundancy for more accuracy.
The motivation is that solutions can still be pruned using entailment or heuristics afterwards.
The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
Duplications should be detected in a post-processing step while accidental constraints should primarily be removed by adding data.

@@

\subsubsection*{Q3. How fast is the algorithm?}
Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$.
The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).

@@

>>>>

We consider various sets of cells in tables:
\begin{itemize}
  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
  \item
  A \textbf{vector} is either a column or a row that is type-consistent.
  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
  \samuel{Vector also has a length; Group has a size}
  \item
  A \textbf{group} is a subrange of vectors with the same orientation in a table.
  We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
  The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
  \item
 A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$. %It defines what it means for a constraint to be well-formed
%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
\end{itemize}
%For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}

@@

\paragraph{Example}
Several example constraints are illustrated in Table 1.  For example, the \template \textit{rank} has
\begin{itemize}
\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are the arguments;
\item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
\samuel{Both must be vectors}
of $v_x$ and $v_r$ be identical, i.e., $length(v_x) = length(v_r)$;
\item \CFunction: \textit{$v_r$ = RANK($v_x$)} whenever each value of $v_r$ is the rank of the corresponding value in $v_x$.
% is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
\end{itemize}

When looking into constraints like \textit{$v_r$ = RANK($v_x$)}, it is important to see the analogy with both first order logic and with constraint satisfaction problems.
From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $v_r$ and $v_x$ are the terms, which can be either variables or values.
This also holds in our setting: when we write
      $\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}$,
we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
\samuel{clarify, e.g. what is group variable}
or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $[1,2,3,4]$. {\bf CHECK}
This also means that we can talk about constraints being satisfied. Basically, a constraint $c(t_1, ..., t_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
if the constraint holds on the values that the $t_i$ take in the set of tables ${\cal T}$. Figure 3 provides a list of constraints that are satisfied by the tables
in Figure 1.

@@

\tias{Redundancy instead of dependency}

@@

\tias{Reformulate: describe there's an naive approach...}
Tables are detected by detecting rectangular collections of data and attempting to detect and separate headers from the data.
Especially this last step is a complex task on its own  \cite{header} and a sophisticated implementation goes beyond the scope of this paper.
On well formated data, however, the user will not have to specify tables manually.

\tias{Split extensions and group detection, clarify group detection, example (e.g. currency for extensions)}
Our algorithm also attempts to detect groups within tables by partitioning the data into consecutive, type consistent groups, essentially separating textual and numerical data.
We support some extensions such as currencies and percentual data and our approach allows us to omit manual specifications of groups.

@@

\tias{Move this to problem statement second part}
A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
We consider $c_1$ to be more specific than $c_2$ or $c_2$ to be more general than $c_1$.
A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.

@@

The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
\\\\
\sergey{Can be parallelized by looking at connected components and respecting order}
Most parts of the algorithm could be easily parallelized.
However, \dependencies would impose a partial order in which constraint templates would have to be treated.

@@

Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
\tias{...such that at least one subgroup satisfies the signature}
For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that those groups or some of their subgroups satisfy the \CSignature.

\tias{Add paragraph on enforcement of group generation}

@@

\subsection{Case study}
\tias{Move explanation of origin to first time it is introduced}
Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
Figure~\ref{fig:sol_example} shows the constraints that we expect to find.

@@

\subsubsection*{Q2. How many redundant constraints are discovered?}
Our primary focus was accuracy, therefore we sometimes traded off less redundancy for more accuracy.
The motivation is that solutions can still be pruned using entailment or heuristics afterwards.
The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
\tias{Advice, move to discussion}
Duplications should be detected in a post-processing step while accidental constraints should primarily be removed by adding data.

@@

\subsubsection*{Q3. How fast is the algorithm?}
\tias{Mention trade-off in the approach}
Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$.
The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).

@@

\sergey{Add summary table of results}
